---
title: "RWorkshop"
output: html_document
date: "2024-09-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Workshop!
Welcome to the R Workshop! This session will cover the basics of R, a powerful programming language used for data analysis and statistical computing. You'll learn how to create variables, work with different data types, and use basic functions.

# Basic R Syntax
```{r}
# Creating variables of different data types
x <- 10                  # Numeric variable
name <- "Sam"          # Character variable
is_student <- TRUE       # Logical variable

# Print the variables
print(x)
print(name)
print(is_student)
```

# Arithmetic Operations in R
You can perform arithmetic operations on numeric data. Let's try some basic operations.
```{r}
# Basic arithmetic operations
sum <- 5 + 3             # Addition
diff <- 10 - 4           # Subtraction
prod <- 2 * 6            # Multiplication
quotient <- 8 / 2        # Division
power <- 2^3             # Exponentiation

# Print the results
sum
diff
prod
quotient
power

```
# Functions in R
R has a variety of built-in functions. You can also create your own functions. Here are some examples:
```{r}
# Using built-in functions
sqrt(25)                 # Square root
log(100)                 # Natural logarithm
round(3.14159, digits=2) # Rounding numbers

# Creating a simple custom function
multiply_by_two <- function(x) {
  return(x * 2)
}

# Using the custom function
multiply_by_two(6)
```
# Working with Vectors
A vector is a basic data structure in R. You can create vectors using the c() function.
```{r}
# Creating a numeric vector
numbers <- c(1, 2, 3, 4, 5)

# Creating a character vector
names <- c("Alice", "Bob", "Charlie")

# Accessing elements of a vector
numbers[1]       # First element
names[2]         # Second element

# Vector arithmetic
numbers * 2      # Multiply each element by 2
```

## Data Analysis
Olympic Summer Games - Paris 2024
Dataset from Kaggle: https://www.kaggle.com/datasets/muhammadehsan02/olympic-summer-games-paris-2024?resource=download

The read.csv() function is commonly used to import CSV files into R.
```{r}
# data <- read.csv("path/to/medallists.csv")
data <- read.csv("/Users/Ramya_1/Desktop/medallists.csv")
```

# Data Structure & Summaries
The str() and summary() functions are used to understand the data structure and statistical summaries.
```{r}
str(data)
summary(data)
```

# Filtering Rows
Use the filter() function from dplyr to extract specific data. For example, filtering by medal_type:
```{r}
library(dplyr) 
gold_medals <- filter(data, medal_type == "Gold Medal")

```

